Operational Performance Analyses at European Airports with Community-Collected Open Data

### 2020-03-10 _paper-2020-ICNS-eidw.Rmd
clean-up after paper submission

@@

# visual inspection - here can ml help?
brg_cut <-c(0,45,115,170,265,360)  # maybe another cut at 325
ggplot(data = asma, mapping=aes(x = BRG, fill=AP_C_RWY)) + 
  geom_histogram(binwidth=5) +
  geom_vline(xintercept=brg_cut, color="red", linetype="dotted") +
  xlim(0,360) +
  theme_minimal() +
  labs(x="bearing from ARP of TMA entry (40NM)") +
  guides(fill=guide_legend(title="arrival runway")) +
  # top right corner
  theme(legend.justification=c(0.9,1), legend.position=c(0.9,1))

ggsave("EIDW-asma-sectors.png", path="./figures")

================================= stopped here ===============================

calc_dist <- function(.trj, .arp){
  df <- .trj %>% mutate(DIST_ARP = geosphere::distHaversine(.arp), cbind(lon,lat))
}

>>>>

if(!exists("eidw_analytic_data")){
  message("trying to load analytica data from file!")
  eidw_analytic_data <- readr::read_csv("./data/EIDW_analytic_data.csv")
}

@@

# visual inspection - here can ml help?
brg_cut <-c(0,45,115,170,265,360)  # maybe another cut at 325
cut_lbls<-c("0-45","45-115","115-170","170-265","265-360")
ggplot(data = asma, mapping=aes(x = BRG, fill=AP_C_RWY)) + 
  geom_histogram(binwidth=5) +
  geom_vline(xintercept=brg_cut, color="red", linetype="dotted") +
  xlim(0,360) +
  theme_minimal() +
  labs(x="bearing from ARP of TMA entry (40NM)") +
  guides(fill=guide_legend(title="arrival runway")) +
  # top right corner
  theme(legend.justification=c(0.9,1), legend.position=c(0.9,1))

# ggsave("EIDW-asma-sectors.png", path="./figures")

# update asma with bearing sectors
asma <- asma %>% 
  mutate( ASMA_SEC = ifelse(point == "NM40", cut(BRG, breaks=brg_cut, labels=cut_lbls), NA)
         ,ASMA_LBL = cut(BRG, breaks=brg_cut)
  )

asma_ind <- asma %>% select(icao24, callsign, DOF, time, AC_CLASS, AP_C_RWY, point, ASMA_SEC, ASMA_LBL) %>%
  group_by(icao24, callsign, DOF) %>% 
  filter(point %in% c("NM40","NM08")) %>% 
  mutate(TRAVEL = ifelse(point == "NM40", lead(time)-time, NA)) %>% 
  filter(point == "NM40") %>% ungroup()

asma_ref <- asma_ind %>% group_by(ASMA_SEC,ASMA_LBL, AP_C_RWY, AC_CLASS) %>%
  summarise(N = n(), N_NA = sum(is.na(TRAVEL)), REF = quantile(TRAVEL, probs=0.2, na.rm=TRUE)) %>%
  ungroup()

asma_ind <- asma_ind %>% inner_join(asma_ref, by=c("ASMA_SEC","ASMA_LBL","AP_C_RWY","AC_CLASS")) %>%
  mutate(ADD_ASMA = TRAVEL - REF) %>%
  summarise(
     ADD_ASMA = sum(ADD_ASMA, na.rm=TRUE), TRAVEL = sum(TRAVEL, na.rm=TRUE)
    ,REF      = sum(REF, na.rm=TRUE)
    ) %>%
  mutate(N = sum(asma_ref$N), N_NA = sum(asma_ref$N_NA), N_VALID = N - N_NA
    ,AVG_ADD_ASMA = ADD_ASMA / N_VALID  
  )

write_csv(asma_ind, "./data/EIDW-study-ASMA.csv")

################# RUN FOR PAPER COMPLETED ##################################

### 2020-03-10 paper01.Rmd
clean-up after paper submission

---
title: "Operational Performance Analyses at European Airports with Community-Collected Open Data"
author: |  # when using the pipe operator do not use any parentheses
  | Antonio Lazarovski, ATOS, Brussels, Belgium 
  | Enrico Spinielli and Rainer Koelle, EUROCONTROL, Brussels, Belgium
output:
  bookdown::pdf_document2:
    toc: false
  bookdown::word_document2:
    toc: false
  html_document:
    df_print: paged
#abstract: My abstract
bibliography: paperxx.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

@@

## Replicating Aerodrome Performance Indicators

@@

The additional taxi-out time is defined as the average of the additional taxi-out times of all departing aircraft.
The additioanl taxi-out time of an individual flight is determined as the difference between the observed actual taxi-out time and an associated reference taxi-out time [@pru_addTXOT_2015].

@@

@@

```{r dips, fig.cap="(ref:dips)", out.width="50%", fig.pos="h", echo=FALSE}
knitr::include_graphics("figures/dips.png")
```

>>>>

---
title: "Operational Performance Analyses at European Airports with Community-Collected Open Data"
author: |  # when using the pipe operator do not use any parentheses
  | Antonio Lazarovski, ATOS, Brussels, Belgium 
  | Enrico Spinielli and Rainer Koelle, EUROCONTROL, Brussels, Belgium
output:
  bookdown::word_document2:
    toc: false
    reference_docx: tmpl-ref.docx
  bookdown::pdf_document2:
    toc: false
  html_document:
    df_print: paged
#abstract: My abstract
bibliography: paperxx.bib
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
```

@@

## Aerodrome Performance Indicators

@@

### Additional Time in Terminal Airspace

The additional time in the terminal airspace is defined as the average of the additional time of arriving traffic observed between the entry within the terminal airspace and landing.
The additional time is determined based on the difference between the observed actual travelling time and a respective reference time for the arrival entry sector, landing runway, and aircraft class, c.f \@ref(eq:addASMA).

\begin{equation}     
(\#eq:addASMA)
avgAddASMA =\frac{\sum_{f_i}{  actualASMA_i - refASMA_{(class,entry,rwy)}  }}{\sum_{f_i}{i}}
\end{equation}

As mentioned above, for Dublin (EIDW) the coverage did not allow to measure landing times.
Accordingly, a proxy for the arrival process **end time** is determined with the 8NM gate.
At this moment in time aircraft are established on final.
It is assumed that the remaining flight time is a systemic time elapsing that is captured by the difference between the travel times and reference times measured with landing timestamps.
Table \@ref(tab:asma-eidw) shows the numerical results for this study and aforementioend approximisation.
For EIDW an average additional ASMA time of 3.64 min/arrival is measured.

(ref:asma-eidw) additional ASMA time measured with OSN data

```{r asma-eidw, echo=FALSE, message=FALSE}
asma_eidw <- readr::read_csv("./data/EIDW-study-ASMA.csv")
knitr::kable(
  asma_eidw, booktabs = TRUE,
  caption = '(ref:asma-eidw)'
)
```

### Additional Taxi-Out Time

The additional taxi-out time is defined as the average of the additional taxi-out times of all departing aircraft.
The additioanl taxi-out time of an individual flight is determined as the difference between the observed actual taxi-out time and an associated reference taxi-out time [@pru_addTXOT_2015].
Taxi out is defined as the time elapsing between the actual off-block and actual take-off of an aircraft.
Accordingly, this metric requires the identification of these surface movement events.
It follows from the previous section that taxi-times cannot be derived from the OSN data feed for Dublin (EIDW) at the time being. This section reports on results for Frankfurt Rhine/Main airport (EDDF)

@@

Table \@ref(tab:txot-eddf) shows the numerical results for this study.
A critical element that requires further analysis is the influence of aircraft for which not stand could had been identified.
This paper applies a grouping approach. Stands located in close proximity are considered as one stand cluster.
Approximately 70% of the traffic was corrently identified in terms of parking positions.
The identification of the used runway was done on the basis of a simple position fix for the take-off event.
With these approximations, an average additional ASMA time of 4.61 min/departure is measured.
This overall results also ranges within the data published by the PRU.

(ref:txot-eddf) additional txot time measured with OSN data

```{r txot-eddf, echo=FALSE, message=FALSE}
txot_eddf <- readr::read_csv("./data/EDDF_analytic_txot.csv")
knitr::kable(
  txot_eddf, booktabs = TRUE,
  caption = '(ref:txot-eddf)'
)
```

@@

```{r dips, fig.cap="(ref:dips)", out.width="50%", fig.pos="h", echo=FALSE}
knitr::include_graphics("./figures/dips-methodology.png")
```
